The hydra and the rooster

### 2020-07-09 latex/chapter-ks.tex
renaming is_limit and is_succ

\begin{Coqsrc}
Lemma canonS_limit_strong (lambda : T1) : 
     nf lambda ->
     is_limit lambda  ->
     forall beta, beta < lambda ->
                  {i:nat | beta < canonS lambda i}.

@@

\begin{Coqsrc}
Lemma canonS_limit_lub (lambda : T1) :
  nf lambda -> is_limit lambda  ->
  strict_lub (fun i => canonS lambda i) lambda.
\end{Coqsrc}

@@

  \begin{Coqsrc}
Theorem Theorem_2_4 (lambda : T1) :
   nf lambda ->
   is_limit lambda  ->
   forall i j, (i < j)%nat ->
               const_pathS 0 (canonS lambda j)
                             (canonS lambda i). 
  \end{Coqsrc}

>>>>

\begin{Coqsrc}
Lemma canonS_limit_strong (lambda : T1) : 
     nf lambda ->
     limitb lambda  ->
     forall beta, beta < lambda ->
                  {i:nat | beta < canonS lambda i}.

@@

\begin{Coqsrc}
Lemma canonS_limit_lub (lambda : T1) :
  nf lambda -> limitb lambda  ->
  strict_lub (fun i => canonS lambda i) lambda.
\end{Coqsrc}

@@

  \begin{Coqsrc}
Theorem Theorem_2_4 (lambda : T1) :
   nf lambda ->
   limitb lambda  ->
   forall i j, (i < j)%nat ->
               const_pathS 0 (canonS lambda j)
                             (canonS lambda i). 
  \end{Coqsrc}

### 2020-07-09 latex/chapter-large-sets.tex
renaming is_limit and is_succ

\begin{Coqsrc}
Section lim.
  Variables (lambda : T1)
            (Hnf : nf lambda)
            (Hlim : is_limit lambda)
            (f : nat -> nat -> nat)
            (H : forall k, L_spec (canonS lambda k) (f (S k))).

@@

Equations  L_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
  L_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Is_Limit alpha) :=
          { | left _ =>  L_ (Canon alpha i)  (S i) ;
            | right notlimit =>  L_ (Pred alpha) (S i)}}.

@@

Lemma L_lim_eqn alpha i : Is_Limit alpha -> L_ alpha i =
                                        L_ (Canon alpha i) (S i).

@@

\begin{Coqsrc}
Equations H_ (alpha: E0) (i:nat) :  nat  by wf  alpha lt :=
  H_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Is_Limit alpha) :=
          { | left _ =>  H_ (Canon alpha (S i))  i ;
            | right notlimit =>  H_ (Pred alpha) (S i)}}.

@@

Lemma H_lim_eqn alpha i : Is_Limit alpha ->
                      H_ alpha i =  H_ (Canon alpha (S i)) i.

@@

\index{Coq!Plug-ins!Equations}
\begin{Coqsrc}
Fail Equations F_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
  F_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Is_Limit alpha) :=
          { | left _ =>  F_ (Canon alpha i)  i ;
            | right notlimit =>  iterate (F_ (Pred alpha))  (S i) i}}.
\end{Coqsrc}

\begin{Coqanswer}
The command has indeed failed with message:
In environment
alpha : E0
notlimit : Is_Limit alpha = false
nonzero : alpha <> Zero
i : nat
F_ : forall x : E0, nat -> x < alpha -> nat
The term "F_ (Pred alpha) ?x" has type "Pred alpha < alpha -> nat"
while it is expected to have type "Pred alpha < alpha -> Pred alpha < alpha"
(cannot unify "nat" and "Pred alpha < alpha").
\end{Coqanswer}

@@

Equations  F_star (c: E0 * nat) (i:nat) :  nat by wf  c call_lt :=
    F_star (alpha, 0) i := i;
    F_star (alpha, 1) i
      with E0_eq_dec alpha Zero :=
           { | left _ => S i ;
             | right nonzero
                 with Utils.dec (Is_Limit alpha) :=
                 { | left _ => F_star (Canon alpha i,1) i ;
                   | right notlimit =>
                     F_star (Pred alpha, S i)  i}};
    F_star (alpha,(S (S n))) i :=
               F_star (alpha, 1) (F_star (alpha, (S n)) i).

@@

Lemma F_lim_eqn : forall alpha i,  Is_Limit alpha ->
                               F_ alpha i = F_ (Canon alpha i) i.

>>>>

\begin{Coqsrc}
Section lim.
  Variables (lambda : T1)
            (Hnf : nf lambda)
            (Hlim : limitb lambda)
            (f : nat -> nat -> nat)
            (H : forall k, L_spec (canonS lambda k) (f (S k))).

@@

Equations  L_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
  L_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Limitb alpha) :=
          { | left _ =>  L_ (Canon alpha i)  (S i) ;
            | right notlimit =>  L_ (Pred alpha) (S i)}}.

@@

Lemma L_lim_eqn alpha i : Limitb alpha -> L_ alpha i =
                                        L_ (Canon alpha i) (S i).

@@

\begin{Coqsrc}
Equations H_ (alpha: E0) (i:nat) :  nat  by wf  alpha lt :=
  H_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Limitb alpha) :=
          { | left _ =>  H_ (Canon alpha (S i))  i ;
            | right notlimit =>  H_ (Pred alpha) (S i)}}.

@@

Lemma H_lim_eqn alpha i : Limitb alpha ->
                      H_ alpha i =  H_ (Canon alpha (S i)) i.

@@

\index{Coq!Plug-ins!Equations}
\begin{Coqsrc}
Fail Equations F_ (alpha: E0) (i:nat) :  nat  by wf  alpha Lt :=
  F_ alpha  i with E0_eq_dec alpha Zero :=
    { | left _ =>  i ;
      | right nonzero
          with Utils.dec (Limitb alpha) :=
          { | left _ =>  F_ (Canon alpha i)  i ;
            | right notlimit =>  iterate (F_ (Pred alpha))  (S i) i}}.
\end{Coqsrc}

\begin{Coqanswer}
The command has indeed failed with message:
In environment
alpha : E0
notlimit : Limitb alpha = false
nonzero : alpha <> Zero
i : nat
F_ : forall x : E0, nat -> x < alpha -> nat
The term "F_ (Pred alpha) ?x" has type "Pred alpha < alpha -> nat"
while it is expected to have type "Pred alpha < alpha -> Pred alpha < alpha"
(cannot unify "nat" and "Pred alpha < alpha").
\end{Coqanswer}

@@

Equations  F_star (c: E0 * nat) (i:nat) :  nat by wf  c call_lt :=
    F_star (alpha, 0) i := i;
    F_star (alpha, 1) i
      with E0_eq_dec alpha Zero :=
           { | left _ => S i ;
             | right nonzero
                 with Utils.dec (Limitb alpha) :=
                 { | left _ => F_star (Canon alpha i,1) i ;
                   | right notlimit =>
                     F_star (Pred alpha, S i)  i}};
    F_star (alpha,(S (S n))) i :=
               F_star (alpha, 1) (F_star (alpha, (S n)) i).

@@

Lemma F_lim_eqn : forall alpha i,  Limitb alpha ->
                               F_ alpha i = F_ (Canon alpha i) i.

### 2020-07-09 latex/chapter-skippable.tex
renaming is_limit and is_succ

This notion is defined generically as a class in
 Module~\href{../src/html/hydras.Prelude.Ordinal-generic.html}{Prelude.Ordinal\_generic} (see Fig~\vref{figure:subsegment}).

@@

\begin{itemize}
\item $\omega^n$ (for some  integer  $n\geq 2$) : the type of $n$-uples of natural numbers, with the lexicographic product of $n$ copies of \texttt{Nat.lt}.
\item  $\omega^\omega$: the set of nonincreasing sequences of natural numbers, lexicographically ordered (also the set of finite multisets of natural numbers).
\item $\epsilon_0$: The set of terms in Cantor normal form (see Chap.~\ref{chap/T1}).
\item $\Gamma_0$: The set of terms in Veblen normal form.
\end{itemize}

>>>>

This notion is defined generically as a class in
 Module~\href{../src/html/hydras.Prelude.Ordinal-generic.html}{Prelude.Ordinal\_generic} see Fig~\vref{figure:subsegment}).

@@

The order on type \texttt{t} is already defined in
Library~\href{https://coq.inria.fr/distrib/current/stdlib/Coq.Relations.Relation_Operators.html}{Relations.Relation\_Operators}.

\begin{Coqanswer}
Inductive
le_AsB (A B : Type) (leA : A -> A -> Prop) (leB : B -> B -> Prop)
  : A + B -> A + B -> Prop :=
    le_aa : forall x y : A, leA x y -> le_AsB A B leA leB (inl x) (inl y)
  | le_ab : forall (x : A) (y : B), le_AsB A B leA leB (inl x) (inr y)
  | le_bb : forall x y : B, leB x y -> le_AsB A B leA leB (inr x) (inr y)
\end{Coqanswer}

\begin{Coqsrc}
Definition lt : relation t := le_AsB _ _ Peano.lt Peano.lt.
Infix "<" := lt : oo_scope.

Definition le := clos_refl _ lt.

Infix "<=" := le : oo_scope.
\end{Coqsrc}

\subsubsection{Finite ordinals as members of \texorpdfstring{$\omega+\omega$}{omega+omega}}

It is easy do define a coercion from type \texttt{nat} into our type \texttt{t}.

\begin{Coqsrc}
Definition fin (n:nat) : t := inl n.
Coercion fin : nat >-> t.
\end{Coqsrc}

The range of \texttt{fin} is called the set of \emph{finite} ordinals.

\index{Coq!Techniques!Coercions} 
\begin{remark}
Beware of coercions and notation scopes!

Let us consider the following goal:

\begin{Coqsrc}
 Goal (6 < 8).
 auto with arith.
\end{Coqsrc}

\begin{Coqanswer}
1 subgoal (ID 9)

  ============================
  6 < 8
\end{Coqanswer}

Please keep in mind that the current notation scope interprets the infix \texttt{``<''} as the predicate \texttt{Omega\_plus\_omega.lt} and not \texttt{Nat.lt}. More,  the coercion mechanism converts the terms \texttt{6:nat} [resp. \texttt{8:nat} ]
into \texttt{inl 6} [resp. \texttt{inl 8}].  So, the initial goal is correctly interpreted by \coq{}, but not as an inequality between two natural numbers.

Anyway, the initial goal is provable, using \texttt{le\_AsB} first constructor.

\begin{Coqsrc}
  constructor; auto with arith.
Qed.
\end{Coqsrc}

\end{remark}
\subsubsection{Definition of \texorpdfstring{$\omega$}{omega}}

We define $\omega$ as the least ordinal which is larger than any finite ordinal.

\begin{Coqsrc}
 Notation  "'omega'"  := (inr 0) : oo_scope. 
\end{Coqsrc}

\begin{exercise}
Show that, for any $\alpha:t$, $\alpha$ is finite iff $\alpha<\omega$.
\end{exercise}

We say also that $\alpha$ is the \emph{limit} of the sequence $\textit{fin}\;i, i\in\mathbb{N}$, according to the followi,g definition (in~\href{../src/html/hydras.Prelude.Ordinal-generic.html}{Prelude.Ordinal\_generic}).
$\alpha$ is the $\omega$-limit of the sequence $s$ if $\alpha$ is [strictly] larger than all the $s(i)$, and for any
$\beta<\alpha$, $\beta$ is dimonated by at least an element of  the sequence $s$ (in fact $\beta$ is dominated by an infinite number of elements of the sequence (\emph{exercise!})).

\begin{Coqsrc}
Definition  Omega_limit
            {A:Type}{lt : relation A}
           {sto : StrictOrder lt} (s: nat -> A) (alpha:A)  :=
  (forall i: nat, lt (s i) alpha) /\
  (forall beta, lt beta  alpha -> exists i:nat, lt beta (s i)).

Lemma omega_is_limit:
   Omega_limit fin omega.
\end{Coqsrc}

\begin{itemize}
\item $\omega^n$ (for some  integer  $n\geq 2$) : the type of $n$-uples of natural numbers, with the lexicographic product of $n$ copies of \texttt{Nat.lt}.
\item  $\omega^\omega$: the set of nonincreasing sequences of natural numbers, lexicographically ordered (also the set of finite multisets of natural numbers).
\item $\epsilon_0$: The set of terms in Cantor normal form (see Chap.~\ref{chap:T1}).
\item $\Gamma_0$: The set of terms in Veblen normal form.
\end{itemize}

### 2020-07-09 latex/chapter-termination.tex
renaming is_limit and is_succ

 \vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#is_succ}{Epsilon0.T1}}

\begin{Coqsrc}
  Fixpoint is_succ alpha :=
  match alpha with
      zero => false
    | ocons zero _ _ => true
    | ocons alpha n beta => is_succ beta
  end.

Fixpoint is_limit alpha :=
  match alpha with
      zero => false
    | ocons zero _ _ => false
    | ocons alpha n zero => true
    | ocons alpha n beta => is_limit beta
  end.
\end{Coqsrc}

\begin{Coqsrc}
  Compute is_limit omega.
\end{Coqsrc}

@@

\begin{Coqsrc}
Compute is_succ 42.
\end{Coqsrc}

@@

\begin{Coqsrc}
  Definition zero_succ_limit (alpha: T1) :
  {is_succ alpha} + {is_limit alpha} +  {alpha=zero}.
  (* ... *)
Defined.
 \end{Coqsrc}

@@

\begin{Coqsrc}
 Lemma is_succ_iff alpha (Halpha : nf alpha) :
  is_succ alpha <-> exists beta : T1, nf beta /\ alpha = succ  beta.
\end{Coqsrc}

>>>>

 \vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Epsilon0.T1.html\#succb}{Epsilon0.T1}}

\begin{Coqsrc}
  Fixpoint succb alpha :=
  match alpha with
      zero => false
    | ocons zero _ _ => true
    | ocons alpha n beta => succb beta
  end.

Fixpoint limitb alpha :=
  match alpha with
      zero => false
    | ocons zero _ _ => false
    | ocons alpha n zero => true
    | ocons alpha n beta => limitb beta
  end.
\end{Coqsrc}

\begin{Coqsrc}
  Compute limitb omega.
\end{Coqsrc}

@@

\begin{Coqsrc}
Compute succb 42.
\end{Coqsrc}

@@

\begin{Coqsrc}
  Definition zero_succ_limit (alpha: T1) :
  {succb alpha} + {limitb alpha} +  {alpha=zero}.
  (* ... *)
Defined.
 \end{Coqsrc}

@@

\begin{Coqsrc}
 Lemma succb_iff alpha (Halpha : nf alpha) :
  succb alpha <-> exists beta : T1, nf beta /\ alpha = succ  beta.
\end{Coqsrc}
